---
title: "Chapter 1"
author: "Marjorie Blanco"
date: "12/19/2018"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
# Clear environment
rm(list = ls(all = TRUE))
knitr::opts_chunk$set(echo = FALSE, message=FALSE, error=FALSE, warning=FALSE)
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

# Clear environment
rm(list = ls(all = TRUE))
library(wooldridge)
library(dplyr)
library(ggplot2)
```


## C1-WAGE1

A data.frame with 526 observations on 24 variables:

- wage: average hourly earnings
- educ: years of education
- exper: years potential experience
- tenure: years with current employer
- nonwhite: =1 if nonwhite
- female: =1 if female
- married: =1 if married
- numdep: number of dependents
- smsa: =1 if live in SMSA
- northcen: =1 if live in north central U.S
- south: =1 if live in southern region
- west: =1 if live in western region
- construc: =1 if work in construc. indus.
- ndurman: =1 if in nondur. manuf. indus.
- trcommpu: =1 if in trans, commun, pub ut
- trade: =1 if in wholesale or retail
- services: =1 if in services indus.
- profserv: =1 if in prof. serv. indus.
- profocc: =1 if in profess. occupation
- clerocc: =1 if in clerical occupation
- servocc: =1 if in service occupation
- lwage: log(wage)
- expersq: exper^2
- tenursq: tenure^2

```{r}
data(WAGE1)
```

- Average education level (years)

```{r}
mean(wage1$educ)
```

- Lowest education level (years)

```{r}
min(wage1$educ)
```

- Highest education level (years)

```{r}
max(wage1$educ)
```
- Average hourly wage

```{r}
mean(wage1$wage)
```

- Women

```{r}
sum(wage1$female)
```
- Men

```{r}
sum(!wage1$female)
```

## C2-BWGHT

A data.frame with 1388 observations on 14 variables:

- faminc: 1988 family income, $1000s
- cigtax: cig. tax in home state, 1988
- cigprice: cig. price in home state, 1988
- bwght: birth weight, ounces
- fatheduc: father’s yrs of educ
- motheduc: mother’s yrs of educ
- parity: birth order of child
- male: =1 if male child
- white: =1 if white
- cigs: cigs smked per day while preg
- lbwght: log of bwght
- bwghtlbs: birth weight, pounds
- packs: packs smked per day while preg
- lfaminc: log(faminc)

```{r}
data('bwght')
```
- Women

```{r}
sum(!bwght$male)
```
- Men

```{r}
sum(bwght$male)
```

```{r}
bwght %>% mutate(smoke = if_else(cigs == 0, 0, 1)) %>% group_by(male) %>% count(smoke)
```

```{r}
#bwght %>% filter(male == 0 & smoke == 1) %>% count()
```

- Average number of cigarettes smoked per day (All)

```{r}
bwght %>% filter(cigs > 0) %>% select(cigs) %>% summary()
ggplot(bwght %>% filter(cigs > 0), aes(x=cigs)) +
  geom_histogram(bins = 50)
```

- Average number of cigarettes smoked per day (women during pregnancy)

```{r}
bwght %>% filter(male == 0 & cigs > 0) %>% select(cigs) %>% summary()
ggplot(bwght %>% filter(male == 0 & cigs > 0), aes(x=cigs)) +
  geom_histogram(bins = 50)
```

```{r}
bwght %>% filter(is.na(fatheduc)) %>% count()  #NAs for fatheduc
bwght %>% filter(!is.na(fatheduc)) %>% count()
mean(bwght$fatheduc, na.rm = TRUE)
```

```{r}
avg_income <- mean(bwght$lfaminc)
avg_income
std_dev_income <- sd(bwght$lfaminc)
std_dev_income
```


## C3-MEAP00_01

A data.frame with 1692 observations on 9 variables:

- dcode: district code
- bcode: building code
- math4: percent students satisfactory, 4th grade math
- read4: percent students satisfactory, 4th grade reading
- lunch: percent students eligible for free or reduced lunch
- enroll: school enrollment
- exppp: expenditures per pupil: expend/enroll
- lenroll: log(enroll)
- lexppp: log(exppp)

```{r}
data('meap00_01')
```

```{r}

```



